Water resource utilization regimes at a basin scale: transition framework and development traps

### 2020-09-29 tex/main/manuscript.tex
update results and refs. to SI figures

%tag 摘要
\begin{abstract}
	From agrarian to industrial societies, humans have harnessed ecosystems over rivers basins around the world, while using water resources as blood that sustains social developments. 
	Although relations between societies and water resources keep changing throughout, there is still lacking effective methods to detect related regimes of water utilization, with much fewer attempts to develop theoretical framework to explain their transitions as well. 
	Here, by integrating three equally important but differentiated dimensions of water utilization (stress, priority and lopsidedness), we develop an Integrated Water Resources Utilization (IWRU) Index at a basin scale to indicate regime shifts. 
	By applying this index to the Yellow River Basin, China, results suggest three water utilization regimes in developing over half a century, whose shifts led by various but pervasive causes.
	Based on that, we summarized a general transition framework which gives a sketch of relationships between human societies and their water utilization, as a useful guideline for big river basins to develop in a coordinated way.
\end{abstract}

@@

% Tag 引言第三段
% 前人已经从不同维度刻画了人水关系.
Development of societies by using water resources has been going on for at least thousands of years. Although its regime shifts and transition phases are not fully understood yet, features of water utilization has been depicted and studied from different perspectives.
% 首先，因为水资源的稀缺性和全球用水量的增加，受到最广泛关注的是人类社会面临的水资源压力。
Firstly, water stresses if of increasing importance and widespread concerns because of scarcity.
Even though the stocks of water in reservoirs are helpful to water resources availability, greater water utilization stresses had become a major constraint to development, because of significant increment in water withdrawals and larger shares of inflexible water utilization during the last century
\cite{postelHumanAppropriationRenewable1996, greveGlobalAssessmentWater2018a, qinFlexibilityIntensityGlobal2019}.
% 其次，随着工业迅猛发展和生态建设的需要，社会对水资源的利用的倾向性也发生了转变。
Secondly, as the need of industry, services and ecology developments, priority of water utilization changed with. 
Despite a major water utilization of agricultural irrigation dominating most river basins, there are noticeable growths and preferential priority in the economy profits and water consumption regarding industry or services, leading potential conflicts between different sectors
\cite{liuWaterScarcityAssessments2017, florkeWaterCompetitionCities2018}.
% 最后，由于水的可用性本质上是区域问题，水资源利用的格局也很重要
Thirdly, since water distribution and utilization are inherently regional concerns where all regions attempt to develop themselves by competitive sectors, lopsidedness also plays as an important aspect.
% 全球总的取水量很少，但缺水地区很多
While only 10\% of available water is withdrawn on global average, about 30\% of population settles in highly water-stressed regions with vary dominated water demands
\cite{wadaWedgeApproachWater2014, okiGlobalHydrologicalCycles2006}.
% 此外，人类活动还在改变这一格局
Human activities further affect this lopsidedness, as positive impacts mostly occur in upper regions whereas aggravated downstream, though most governments are trying to control this situation. 
\cite{veldkampWaterScarcityHotspots2017}. 
% 总的来说，上述三个维度试图解答与水资源利用有关的三个重要问题 “还有多少水资源可以利用”，“把水资源用在什么”
Taken together, existing researches have evaluated the three dimensions of water resource utilization regarding a series of crucial questions: “How much water resources?”, “How to use them?” and “Used for whom or in where?”.
% 将三种视角结合，就是“水资源利用体系”。
However, as these dimensions haven't been well integrated by quantitative methods, there is still lacking a coherent interpret of regime shifts regard to social development and water utilization.

@@

\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/framework.jpg}
	\caption{
		A framework for understanding the changing relationship between watershed development and water resources.
		% 图A是水资源利用的三个维度。每个维度都有两个极点（红色字表示），指示水资源利用在该轴上的两个变化方向。
		\textbf{A,} three dimensions (stress, priority and lopsidedness) of water resources utilization. Each dimension has two poles (denoted in red) which indicates the two potential directions of water resources utilization changes along that axis. (1) Stress of water use shifts between low and high, which means there is no shortage of water supply or there is. (2) Priority of water utilization can move between a provisioning part or a non-provisioning one, indicating how much water were used in food supporting to human societies. (3) Lopsidedness can move between balanced or lopsided, when allocation of water resources between different sectors or regions changed. We presume that water utilization regimes equally weighted by these basic dimensions, whose combination can highly relate to basin development. 
		% 图B是将三个维度结合后的变化情况。因上述三个维度随着社会发展而不断变化，其组合的水资源利用状态也不同。这个过程中当突变发生时，可能标志着水资源利用发生了稳态转换，因此我们需要一个指标来监测其变化。
		\textbf{B,} the changes after combining the three dimensions. Because the above three dimensions are changing with the development of society, their combined water resource utilization status is also different. When abrupt transitions occur during this process, they may indicate a regime shift in water utilization, so we need an indicator to monitor this change.
	}
	\label{fig:framework}
\end{figure*}

\section*{Results}
\subsection*{Detection of Water utilization regimes}
% 这一节主要展示IWRU的变化趋势和WUR的划分
With two significant points, the trend of IWRU index are detected into three periods, whose slopes of changing are various and mainly contributed by different dimensions (stress, priority or lopsidedness of water utilization, see Methods.) (Figure~\ref{fig:IWRU}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/index.pdf}
	\caption{Changes of the IWRU index. 
	\textbf{A,} with two change points in 1978 and 1994, three periods were detected in changing trend of the IWRU.
	\textbf{B,} changes of IWRU in three periods have various slopes, while the second period have a negative growths rate.
	\textbf{C,} changes of the IWRU within three certain periods, which have different main contributors.
	}
	\label{fig:IWRU}
\end{figure}
% 接下来分句介绍每个阶段的特征
% 第一阶段
In the first period (P1, 1965-1978), the IWRU index had a rapidly increasing and the lightening of water stresses made the most striking contribution (+0.722 change contribution and 83.7\% net contribution), while priority and lopsidedness of the water utilization had slight negative contribution (-0.048, 5.6\% and -0.09, 10.7\% respectively).
% 第二阶段
In the second period (P2, 1979-1994), the IWRU index experienced a slight drop, despite positive contributions of priority and lopsidedness of water utilization (+0.352 and +0.279 in change contributions, 22.0\% and 27.8\% in net contributions respectively), because of increasing stresses on water resource playing a larger negative role (-0.636, 50.2\%). 
% 第三阶段
However, as the further increasing of positive contributions of water utilization priority (+0.485 change contribution and 36.1\% net contribution) and lopsidedness (+0.515 and 38.3\%, corresponding), and decelerations of water stresses (-0.344 change contribution, 46\% less than P2) in the third period (P3, 1995-2013), a positive growth of the IWRU returned.
% 总而言之，每个阶段都由水资源利用的不同维度提供最大的正向作用
As a result, each period has a different most striking positive contributor to IWRU: P1 is stress; P2 is priority; and P3 is lopsidedness.

% 用水三个维度的组合呈现出阶段特征明显
Combining these three dimensions' net contribution (differ from change contributions, see Method) to IWRU further, ratios of the contributions of the three dimensions clustered clearly by different time periods, indicating three regimes (Figure~\ref{fig:phases}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/phases.pdf}
	\caption{Combination of net contributions regards three dimensions in different periods (S: stresses; P: priority; L: lopsidedness). The closer a point to an angle, greater the proportion of the net contribution of this dimension.
	The red indicator line in this ternary plot denotes 1:1 contributions between priority (P) and lopsidedness (L). When the points are bellow this line, the net contribution ratio of lopsidedness is lower than that of priority, and vice versa.}
	% 由于阶段一的点位于该线上方，L的净贡献比例多于P，而第二阶段的点则恰好相反。
	\label{fig:phases}
\end{figure}
% 第一阶段到第二阶段
At the very beginning (1965) and throughout the whole P1, water utilization regime dominated by high stress. After then, it experienced a shift to low stress since 1978, with a change in the proportion of net contributions between priority and lopsidedness, too.
% 第三阶段集中
Finally, the net contribution of three dimensions were much similar in P3, making the points highly concentrated at the centre of the ternary diagram in that period.

%tag 结果2
\subsection*{Differences between water utilization regimes}
% 在不同的维度下各阶段进行对比，可以发现不同的水资源利用体系间存在明显差异。
The differences between the water utilization regimes are reflected in changes of all three dimensions.
\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/dimensions.pdf}
	\caption{
		Changes in different dimensions of water resources utilization regimes and their main contributors.
		\textbf{A,} changes of water utilization stress, indicated by unstandardized scarcity-flexibility-variability water stresses index (SFV-index).
		\textbf{B,} changes of water utilization priority, indicated by non-provisioning water shares.
		\textbf{C,} changes of water utilization lopsidedness, indicated by unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		\textbf{D,} Main impact factors to water utilization stresses in each period or region, and their contributions to changes of unstandardized SFV-index.
		\textbf{E,} Main impact water uses to water utilization tendentiousnesses, and their contributions to changes of non-provisioning water shares.
		\textbf{F,} Main impact factors to water utilization patterns, and their contributions to changes of related unstandardized distribution information entropy index (see supplementary: Methods. S3 for details).
		}
	\label{fig:dimensions}
\end{figure*}
% 从阶段一到阶段二，水资源压力的变化最为明显
Moving from the regime in P1 to P2, the most striking change is the reversal of the trend in water utilization stress, which is determined by a combination of scarcity, flexibility and variability (Figure~\ref{fig:dimensions}A and Figure~\ref{fig:dimensions}D).   
% 1. 水资源丰富，耗水量少且灵活用水
In the P1, natural surface water resources were rather abundant with lesser water consumptions (supplementary Fig. S1 and S3) and most of which were flexible water utilization (supplementary Fig. S4). During the P1 and even P2, however, water consumption increases rapidly and natural surface water resources decreases at the same time, making water increasingly scarce. Opposite effect to that, numerous reservoirs built reduced the variability of water resources by boosting storage capacities, but most of which built in P1 (supplementary Fig. S5 and supplementary Table S1). As a result, water utilization stress decreases during P1, but begins to rise rapidly in P2.

% 另一方面，P2-P3，持续变化的水资源利用倾向与格局
On the other hand, as the most positive contributors to the IWRU index in P2 and P3 separately, priority and patterns of water utilization were keeping to enlarge their impacts (Figure~\ref{fig:dimensions}B and Figure~\ref{fig:dimensions}C). 
% 首先是用水比例的变化
Representing priority of water utilization, increasing non-provisioning share of water utilization were mainly contributed by larger industrial water consumptions and minor total water uses, while the influence of both are weakening (Figure~\ref{fig:dimensions}E).
% 再讨论用水格局的变化
However, patterns of water utilization, whose contributions to the IWRU are increasing, were mainly benefited from decreasing disparities in the amount of water resources used, both intersectoral and regional (Figure~\ref{fig:dimensions}F).

% tag 结果3
\subsection*{Causes of water utilization regime changes}

\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/causes.pdf}
	\caption{
		Causes of water utilization regime shifts: economy growths, efficiency changes, and managements.
		\textbf{A.} Changes of total irrigated area and water consumption per unit of irrigated area.
		\textbf{B.} Changes of gross values added (GVA) of industry and services, and their water use density (WUI) respectively (see supplementary: Methods. S2 for details).
		\textbf{C.} Constructions' finished time of each new reservoir and their located regions' water use percentages in basin's total water use, at that time. Red ones denote hub reservoirs in the basin, which plays a role in integrated water management. Size of the points indicates their magnitude of water storage capacities. Some important or special reservoirs' name are denoted: (1) Xiaolangdi reservoir and Sanmen Reservoir were constructed mainly responsible for managing sediments of the Yellow River. (2) Liujia Gorge, Longyang Gorge, were constructed mainly responsible for managing water flood discharge and storage. Therefore, these marked reservoirs are significant for the entire basin, far crucial than regional development.
	}
	\label{fig:Causes}
\end{figure}

Some main drivers caused the above changes of water utilization regime.
% 经济总量提升导致资源耗竭的加速。
(1) The expansion of irrigated area and the economic growth of industry and services are keys to the changes in the priority of water utilization between P1 and P2 (Figure\ref{fig:Causes} A). During the P1, irrigated agricultural area in the Yellow River basin expanded rapidly at a rate of $0.25*10^6/year$, and irrigation water was the dominant utilization way ($81.56\%$ of the total water use in 1965, and $83.17\%$ in 1978). Entering P2, however, while the expansion of irrigated area stalled, industry and services gradually took off and took up more water resources (Figure\ref{fig:Causes} B), leading to $8\%$ reduction of proportion of irrigation water.

% 用水关系的变化
(2) During the P3, irrigation, whose water consumptions were still dominating, have noticeable changes in its efficiency, however. Although irrigated agricultural area resumed expansion, and both industry, urban services were boosting their gross added values (GVA), water use density (WUI) experienced significant declines and reached the lowest points (Figure\ref{fig:Causes} A and B). It means, water utilization ways  have changed, along with technological solutions and a range of water conservation practices (supplementary Fig. S6). As a result, the proportions of water use between the different sectors tend to average out while the total water consumption remains stable, after the P3 (supplementary Fig. S7).

% 不断变化的管理模式。
(3) Changing water management practice contributed throughout all three periods. In the P1, most of the reservoirs are built in regions with high water demands, as ratio of regional water use and basinal water use for each new reservoir are significantly higher (Figure~\ref{fig:Causes}C). In the P2, on the other hand, the number of new reservoirs decreases significantly with little increment of total storage capacities (supplementary Fig. S5). Entering the P3, however, the number of new reservoirs are even much higher than that in the P1, and most of them were built in regions with lower ratio of regional water use and basinal water use (Figure~\ref{fig:Causes}C and supplementary Fig. S5).

@@

	\subsection*{Integrated Water Resources Utilization (IWRU) Index}
	The Integrated Water Resources Utilization (IWRU) Index consists of three dimensions (stresses, tendentiousnesses and patterns, denoted by sub-indicators correspondingly). Assuming they have equal weights, we log-transformed and standardized the three sub-indicators for elimination of differences between indicators. It means for each indicator $I_{sub}$, we performed:
	$$ I' = log(I_{sub}) $$
	$$ I = (I' - I'_{min}) / (I'_{max}-I'_{min})$$
	where $I$ is standardized series for $I_{sub}$.
	Then, since we assumed different relationships between development and sub-indicators, we added them together in equal weights:
	$$ IWRU = \sum_{i}^3 I'_i $$
	where $i$ is stress, priority or lopsidedness, and $I'_i$ is standardized sub-indicator of them. A brief description of the sub-indicators used to measure each dimension follows (see SI Methods. S3 for more details).

@@

	% 计算贡献度
	\subsection*{Contribution decomposition}
	\subsubsection*{Change contributions}
		We have decomposed the amount of variation in each index at different stages in order to observe the contribution of each influencing factor to them. Use Integrated Water Resources Utilization (IWRU) Index as an example, which influenced by three dimensions: stress ($S$), priority ($T$) and lopsidedness ($P$) (indicated by their own index respectively, see Water utilization regime index and supplementary Methods. S3):
		$$ IWRU = T * P * S ^ {-1} $$
		Take the logarithm of both sides then, we get:
		$$ ln(IWRU) = ln(S) + ln(T) - ln(P) $$
		Since the changes of IWRU $\Delta IWRU$ can be expressed as $\Delta IWRU = ln(IWRU_2) - ln(IWRU_1)$, where $IWRU_2$ and $IWRU_1$ are ending and beginning values of IWRU in a certain time period, combining the above equations we can get:
		$$ \Delta IWRU = ln(\frac{S_1}{S_2}) + ln(\frac{T_2}{T_1}) + ln(\frac{P_2}{P_1}) = -\Delta S + \Delta T + \Delta P $$
		Then, we can calculate contributions $C_F$ of a certain factor $F$ in a certain period by:
		$$ C_{F} = \frac{|\Delta C_{F}|}{\Delta IWRU}$$

>>>>

%tag 摘要
\begin{abstract}
	From agrarian to industrial societies, humans have harnessed ecosystems over rivers basins around the world, while using water resources as blood that sustains social developments. 
	Although relations between societies and water resources keep changing throughout, there is still lacking effective methods to detect related regimes of water utilization, with much fewer attempts to develop theoretical framework to explain their transitions as well. 
	Here, by integrating three equally important but differentiated dimensions of water utilization (stress, priority and configuration), we develop an Integrated Water Resources Utilization (IWRU) Index at a basin scale to indicate regime shifts. 
	By applying this index to the Yellow River Basin, China, results suggest three water utilization regimes in developing over half a century, whose shifts led by various but pervasive causes.
	Based on that, we summarized a general transition framework which gives a sketch of relationships between human societies and their water utilization, as a useful guideline for big river basins to develop in a coordinated way.
\end{abstract}

@@

% Tag 引言第三段
% 前人已经从不同维度刻画了人水关系.
Development of societies by using water resources has been going on for at least thousands of years. Although its regime shifts and transition phases are not fully understood yet, features of water utilization has been depicted and studied from different perspectives.
% 首先，因为水资源的稀缺性和全球用水量的增加，受到最广泛关注的是人类社会面临的水资源压力。
Firstly, water stresses if of increasing importance and widespread concerns because of scarcity.
Even though the stocks of water in reservoirs are helpful to water resources availability, greater water utilization stresses had become a major constraint to development, because of significant increment in water withdrawals and larger shares of inflexible water utilization during the last century
\cite{postelHumanAppropriationRenewable1996, greveGlobalAssessmentWater2018a, qinFlexibilityIntensityGlobal2019}.
% 其次，随着工业迅猛发展和生态建设的需要，社会对水资源的利用的倾向性也发生了转变。
Secondly, as the need of industry, services and ecology developments, priority of water utilization changed with. 
Despite a major water utilization of agricultural irrigation dominating most river basins, there are noticeable growths and preferential priority in the economy profits and water consumption regarding industry or services, leading potential conflicts between different sectors
\cite{liuWaterScarcityAssessments2017, florkeWaterCompetitionCities2018}.
% 最后，由于水的可用性本质上是区域问题，水资源利用的格局也很重要
Thirdly, since water distribution and utilization are inherently regional concerns where all regions attempt to develop themselves by competitive sectors, configuration also plays as an important aspect.
% 全球总的取水量很少，但缺水地区很多
While only 10\% of available water is withdrawn on global average, about 30\% of population settles in highly water-stressed regions with vary dominated water demands
\cite{wadaWedgeApproachWater2014, okiGlobalHydrologicalCycles2006}.
% 此外，人类活动还在改变这一格局
Human activities further affect this configuration, as positive impacts mostly occur in upper regions whereas aggravated downstream, though most governments are trying to control this situation. 
\cite{veldkampWaterScarcityHotspots2017}. 
% 总的来说，上述三个维度试图解答与水资源利用有关的三个重要问题 “还有多少水资源可以利用”，“把水资源用在什么”
Taken together, existing researches have evaluated the three dimensions of water resource utilization regarding a series of crucial questions: “How much water resources?”, “How to use them?” and “Used for whom or in where?”.
% 将三种视角结合，就是“水资源利用体系”。
However, as these dimensions haven't been well integrated by quantitative methods, there is still lacking a coherent interpret of regime shifts regard to social development and water utilization.

@@

\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/framework.jpg}
	\caption{
		A framework for understanding the changing relationship between watershed development and water resources.
		% 图A是水资源利用的三个维度。每个维度都有两个极点（红色字表示），指示水资源利用在该轴上的两个变化方向。
		\textbf{A,} three dimensions (stress, priority and configuration) of water resources utilization. Each dimension has two poles (denoted in red) which indicates the two potential directions of water resources utilization changes along that axis. (1) Stress of water use shifts between low and high, which means there is no shortage of water supply or there is. (2) Priority of water utilization can move between a provisioning part or a non-provisioning one, indicating how much water were used in food supporting to human societies. (3) configuration can move between balanced or lopsided, when allocation of water resources between different sectors or regions changed. We presume that water utilization regimes equally weighted by these basic dimensions, whose combination can highly relate to basin development. 
		% 图B是将三个维度结合后的变化情况。因上述三个维度随着社会发展而不断变化，其组合的水资源利用状态也不同。这个过程中当突变发生时，可能标志着水资源利用发生了稳态转换，因此我们需要一个指标来监测其变化。
		\textbf{B,} the changes after combining the three dimensions. Because the above three dimensions are changing with the development of society, their combined water resource utilization status is also different. When abrupt transitions occur during this process, they may indicate a regime shift in water utilization, so we need an indicator to monitor this change.
	}
	\label{fig:framework}
\end{figure*}

\section*{Results}
\subsection*{Detection of Water utilization regimes}
% 这一节主要展示IWRU的变化趋势和WUR的划分
With two significant points, the trend of IWRU index are detected into three periods, whose slopes of changing are various and mainly contributed by different dimensions (stress, priority or configuration of water utilization, see \textit{Methods}) (Figure~\ref{fig:IWRU}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/index.pdf}
	\caption{Changes of the IWRU index. 
	\textbf{A,} with two change points in 1978 and 1994, three periods were detected in changing trend of the IWRU.
	\textbf{B,} changes of IWRU in three periods have various slopes, while the second period have a negative growths rate.
	\textbf{C,} changes of the IWRU within three certain periods, which have different main contributors.
	}
	\label{fig:IWRU}
\end{figure}
% 接下来分句介绍每个阶段的特征
% 第一阶段
In the first period (P1, 1965-1978), the IWRU index had a rapidly increasing and the lightening of water stresses made the most striking contribution (+0.722 change contribution and 83.7\% net contribution), while priority and configuration of the water utilization had slight negative contribution (-0.048, 5.6\% and -0.09, 10.7\% respectively).
% 第二阶段
In the second period (P2, 1979-1994), the IWRU index experienced a slight drop, despite positive contributions of priority and configuration of water utilization (+0.352 and +0.279 in change contributions, 22.0\% and 27.8\% in net contributions respectively), because of increasing stresses on water resource playing a larger negative role (-0.636, 50.2\%). 
% 第三阶段
However, as the further increasing of positive contributions of water utilization priority (+0.485 change contribution and 36.1\% net contribution) and configuration (+0.515 and 38.3\%, corresponding), and decelerations of water stresses (-0.344 change contribution, 46\% less than P2) in the third period (P3, 1995-2013), a positive growth of the IWRU returned.
% 总而言之，每个阶段都由水资源利用的不同维度提供最大的正向作用
As a result, each period has a different most striking positive contributor to IWRU: P1 is stress; P2 is priority; and P3 is configuration.

% 用水三个维度的组合呈现出阶段特征明显
Combining these three dimensions' net contribution (differ from change contributions, see Method) to IWRU further, ratios of the contributions of the three dimensions clustered clearly by different time periods, indicating three regimes (Figure~\ref{fig:phases}).
\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/phases.pdf}
	\caption{Combination of net contributions regards three dimensions in different periods (S: stresses; P: priority; L: configuration). The closer a point to an angle, greater the proportion of the net contribution of this dimension.
	The red indicator line in this ternary plot denotes 1:1 contributions between priority (P) and configuration (L). When the points are bellow this line, the net contribution ratio of configuration is lower than that of priority, and vice versa.}
	% 由于阶段一的点位于该线上方，L的净贡献比例多于P，而第二阶段的点则恰好相反。
	\label{fig:phases}
\end{figure}
% 第一阶段到第二阶段
At the very beginning (1965) and throughout the whole P1, water utilization regime dominated by high stress. After then, it experienced a shift to low stress since 1978, with a change in the proportion of net contributions between priority and configuration, too.
% 第三阶段集中
Finally, the net contribution of three dimensions were much similar in P3, making the points highly concentrated at the centre of the ternary diagram in that period.

%tag 结果2
\subsection*{Differences between water utilization regimes}
% 在不同的维度下各阶段进行对比，可以发现不同的水资源利用体系间存在明显差异。
The differences between the water utilization regimes are reflected in changes of all three dimensions.
\begin{figure*}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/dimensions.pdf}
	\caption{
		Changes in different dimensions of water resources utilization regimes and their main contributors.
		\textbf{A,} changes of water utilization stress, indicated by unstandardized scarcity-flexibility-variability water stresses index (SFV-index, see \textit{Methods} and \textit{SI Appendix} Method S4).
		\textbf{B,} changes of water utilization priority, indicated by non-provisioning water shares (see Methods S4).
		\textbf{C,} changes of water utilization configuration, indicated by unstandardized distribution information entropy index (\textit{Methods} and \textit{SI Appendix} Method S4).
		\textbf{D,} Main impact factors to water utilization stresses in each period or region, and their change contributions to unstandardized SFV-index.
		\textbf{E,} Main impact water uses to water utilization tendentiousnesses, and their change contributions to non-provisioning water shares.
		\textbf{F,} Main impact factors to water utilization patterns, and their change contributions of related unstandardized distribution information entropy index (\textit{Methods} and \textit{SI Appendix} Method S4).
		}
	\label{fig:dimensions}
\end{figure*}
% 从阶段一到阶段二，水资源压力的变化最为明显
Moving from the regime in P1 to P2, the most striking change is the reversal of the trend in water utilization stress(Figure~\ref{fig:dimensions}A), which is determined by a combination of scarcity, flexibility and variability (Figure~\ref{fig:dimensions}D).   
% 1. 水资源丰富，耗水量少且灵活用水
In the P1, natural surface water resources were rather abundant with fewer water consumptions (\textit{SI Appendix} Fig. S3) and most of which were flexible water utilization (\textit{SI Appendix} Fig. S4). During the P1 and even P2, however, water consumption increases rapidly and natural surface water resources decreases at the same time, making water increasingly scarce. Opposite effect to that, numerous reservoirs built reduced the variability of water resources by boosting storage capacities, but there are much fewer reservoirs built in P2 (\textit{SI Appendix S5}). As a result, water utilization stress decreases during P1, but begins to rise rapidly in P2.

% 另一方面，P2-P3，持续变化的水资源利用倾向与格局
On the other hand, as the most positive contributors to the IWRU index in P2 and P3 separately, priority (Figure~\ref{fig:dimensions}B) and configuration (Figure~\ref{fig:dimensions}C) of water utilization were keeping to enlarge their impacts. 
% 首先是用水比例的变化
Representing priority of water utilization, increasing non-provisioning share of water utilization were mainly contributed by larger industrial water consumptions and minor total water uses, while their influences are weakening (Figure~\ref{fig:dimensions}E).
% 再讨论用水格局的变化
However, configuration of water utilization, whose contributions to the IWRU are increasing, were mainly benefited from decreasing disparities in the amount of water resources used, both intersectoral and regional (Figure~\ref{fig:dimensions}F).

% tag 结果3
\subsection*{Drivers of water utilization regime shifts}

\begin{figure}%[htbp]
	\centering
	\includegraphics[width=\linewidth]{../../figures/main/causes.pdf}
	\caption{
		Drivers of water utilization regime shifts: economy growths, efficiency changes, and management practices.
		\textbf{A.} Changes of total irrigated area, and water consumptions in per unit of area (\textit{SI Method} S2).
		\textbf{B.} Changes of gross values added (GVA) of industry and services, and their water use density (WUI) respectively (\textit{SI Method} S2).
		\textbf{C.} Constructions' finished time of each new reservoir and their located regions' water use percentages in basin's total water use, at that time. Red ones denote hub reservoirs in the basin, which plays a role in bainsal integrated water management. Size of the points indicates their magnitude of water storage capacities. Some important or special reservoirs' name denoted: (1) Xiaolangdi reservoir and Sanmen Reservoir were constructed mainly responsible for managing sediments of the Yellow River. (2) Liujia Gorge, Longyang Gorge, were constructed mainly responsible for managing water flood discharge and storage. These marked reservoirs, therefore are significant for the entire basin, far crucial than regional development.
	}
	\label{fig:Causes}
\end{figure}

Some main drivers caused the above changes of water utilization regime.
% 经济总量提升导致资源耗竭的加速。
(1) The expansion of irrigated area and the economic growth of industry and services are keys to the changes in the priority of water utilization between P1 and P2 (Figure\ref{fig:Causes} A). During the P1, irrigated agricultural area in the Yellow River basin expanded rapidly at a rate of $0.25*10^6/year$, and irrigation water was the dominant utilization way ($81.56\%$ of the total water use in 1965, and $83.17\%$ in 1978, see \textit{SI Appendix}). Entering P2, however, while the expansion of irrigated area stalled, industry and services gradually took off and took up more water resources (Figure\ref{fig:Causes} B), leading to $8\%$ reduction of proportion of irrigation water (\textit{SI Appendix}).

% 用水关系的变化
(2) During the P3, irrigation had noticeable changes in its efficiency, whose water consumptions were reduced but still dominant. Although irrigated agricultural area resumed expansion, and both industry, urban services were boosting their gross added values (GVA), water use density (WUI) experienced significant declines and reached the lowest points (Figure\ref{fig:Causes} A and B). It means, water utilization ways  have changed, along with technological solutions and a range of water conservation practices (\textit{SI Appendix} Fig. S7). As a result, the differences between the sectors of water use reduced while the total water consumption remains stable, during the P3 (\textit{SI Appendix} Fig. S6).

% 不断变化的管理模式。
(3) Changing water management practice contributed throughout all three periods. In the P1, most of the reservoirs are built in regions with high water demands, as ratio of regional water use and basinal water use for each new reservoir are significantly higher (Figure~\ref{fig:Causes}C, p<0.01). In the P2, on the other hand, the number of new reservoirs decreases significantly with little increment of total storage capacities (\textit{SI Appendix} Fig. S5). Entering the P3, however, the number of new reservoirs are even much higher than that in the P1, and most of them were built in regions with lower ratio of regional water use and basinal water use (Figure~\ref{fig:Causes}C and \textit{SI Appendix} Fig. S5).

@@

	\subsection*{Integrated Water Resources Utilization (IWRU) Index}
	The Integrated Water Resources Utilization (IWRU) Index consists of three dimensions (stresses, tendentiousnesses and patterns, denoted by sub-indicators correspondingly). Assuming they have equal weights, we log-transformed and standardized the three sub-indicators for elimination of differences between indicators. It means for each indicator $I_{sub}$, we performed:
	$$ I' = log(I_{sub}) $$
	$$ I = (I' - I'_{min}) / (I'_{max}-I'_{min})$$
	where $I$ is standardized series for $I_{sub}$.
	Then, since we assumed different relationships between development and sub-indicators, we added them together in equal weights:
	$$ IWRU = \sum_{i}^3 I'_i $$
	where $i$ is stress, priority or configuration, and $I'_i$ is standardized sub-indicator of them. A brief description of the sub-indicators used to measure each dimension follows (see SI Methods. S3 for more details).

@@

	% 计算贡献度
	\subsection*{Contribution decomposition}
	\subsubsection*{Change contributions}
		We have decomposed the amount of variation in each index at different stages in order to observe the contribution of each influencing factor to them. Use Integrated Water Resources Utilization (IWRU) Index as an example, which influenced by three dimensions: stress ($S$), priority ($T$) and configuration ($P$) (indicated by their own index respectively, see Water utilization regime index and supplementary Methods. S3):
		$$ IWRU = T * P * S ^ {-1} $$
		Take the logarithm of both sides then, we get:
		$$ ln(IWRU) = ln(S) + ln(T) - ln(P) $$
		Since the changes of IWRU $\Delta IWRU$ can be expressed as $\Delta IWRU = ln(IWRU_2) - ln(IWRU_1)$, where $IWRU_2$ and $IWRU_1$ are ending and beginning values of IWRU in a certain time period, combining the above equations we can get:
		$$ \Delta IWRU = ln(\frac{S_1}{S_2}) + ln(\frac{T_2}{T_1}) + ln(\frac{P_2}{P_1}) = -\Delta S + \Delta T + \Delta P $$
		Then, we can calculate contributions $C_F$ of a certain factor $F$ in a certain period by:
		$$ C_{F} = \frac{|\Delta C_{F}|}{\Delta IWRU}$$

### 2020-09-29 tex/sup/supplementary_information.tex
update results and refs. to SI figures

% 补充图片3：天然水资源量
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../../figures/sup/sf_measured_runoff.pdf}
    \caption{Natural measured runoff of Yellow River within different periods.}
\end{figure}

% 补充图片4：水资源灵活性
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/inflexible_wu.pdf}
    \caption{Flexibility}
\end{figure}

% 补充图片5：水库数量和累积库容
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/reservoirs.pdf}
    \caption{Reservoirs and accumulated storage}
\end{figure}

% 补充图片6：节水措施
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/saving_water.jpg}
    \caption{technological solutions and water conservation practices}
\end{figure}

% 补充图片7：用水比例
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../../figures/sup/sf_wu_sections_stackplot.jpg}
    \caption{Proportions of water use between the different sectors}
\end{figure}

>>>>

% 补充图片3：天然水资源量
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{../../figures/sup/sf_measured_runoff.pdf}
    \caption{Water resources in different regions. 
        \textbf{A,} changing trend of measured runoff, \textbf{B, C and D} average measured runoff and water consumptions within different periods.
    % TODO 这里增加一个耗水量的图
    }
\end{figure}

% 补充图片4：水资源灵活性
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/inflexible_wu.pdf}
    \caption{
        Flexibility and its change contributions of water utilization.
        \textbf{A,} Proportion of inflexible water uses.
        \textbf{B,} Change contributions of inflexible water uses.
    }
\end{figure}

% 补充图片5：水库数量和累积库容
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/reservoirs.pdf}
    \caption{
        Reservoirs and their storage.
        \textbf{A,} number of new reservoirs in different periods.
        \textbf{B,} total storage capacities in different regions. 
    }
\end{figure}

% 补充图片6：用水比例
% TODO 这里只保留堆积图即可，做的更好看点
\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{../../figures/sup/sf_wu_sections_stackplot.jpg}
    \caption{Proportions of total water use between the different sectors}
\end{figure}

% 补充图片7：节水措施
% TODO 这里增加更丰富的节水措施图
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{../../figures/sup/saving_water.jpg}
    \caption{
        technological solutions and water conservation practices
    }
\end{figure}
